---
title: "Assignment 1a (Python) - refactor obesity dataset code"
output: html_notebook
---

The below code downloads data from the obesity dataset from the [UC Irvine ML repository](https://doi.org/10.24432/C5H31Z). It then preprocesses the data, fits a machine-learning model and shows results. The coding style intentionally is not perfect.

In this assignment you'll refactor this code using the principles taught in the Reproducible Coding practices course. Things to consider are for example:

* Modularity
* Documentation
* Tests
* Packaging
* Dependencies
* Version control
* etc.

For the full assessment criteria see the Rubric. For full details on the assignment see the `README.md`

```{r}
install.packages('httr', repos='http://cran.us.r-project.org') #I think it is better to use the 'renv' function for this packages so that they will have their own private package library.
install.packages('readr', repos='http://cran.us.r-project.org')
install.packages('ggplot2', repos='http://cran.us.r-project.org')
install.packages('randomForest', repos='http://cran.us.r-project.org')

```
```{r}
####Install renv package####
install.packages('renv')
renv::init()
renv::snapshot()
renv::status()
```

```{r}
library(httr)
library(readr)
library(ggplot2)
library(randomForest)
```

```{r}
url <- 'https://archive.ics.uci.edu/static/public/544/estimation+of+obesity+levels+based+on+eating+habits+and+physical+condition.zip'

r <- GET(url)

tmp <- GET(url, write_disk('data.zip', overwrite = TRUE))
stop_for_status(tmp)

unzip('data.zip', exdir = './data')

df <- read_csv('data/ObesityDataSet_raw_and_data_sinthetic.csv', show_col_types = FALSE)
```

####Keep this part, since it is used to open the .zip file.####

```{r}
df_web <- df[1:498, ]
df_synthetic <- df[499:nrow(df), ]
df_web
```

```{r}

####Split data function####
split_data <-function (data, cutoff = 498) {
  list(web = data[1:cutoff, ], synthetic = data[(cutoff + 1):nrow(data), ])
}
```
#allows cutoff value to be chnged describes the function and makes it more modular and reusable.
```{r}
X_columns <- c('Gender', 'Age', 'BMI')

df_web$BMI <- df_web$Weight / (df_web$Height)^2
df_synthetic$BMI <- df_synthetic$Weight / (df_synthetic$Height)^2
```

```{r}
####Calculate BMI used function####
calculate_bmi <- function(weight, height) {
  return(weight / (height^2))
}  # Those codes are repeated, so I think it is beter to use a function to shorten the code and make it more modular. 
df_web$BMI <- calculate_bmi(df_web$Weight, df_web$Height)
df_synthetic$BMI <- calculate_bmi(df_synthetic$Weight, df_synthetic$

plot(df_web$BMI, type = 'l')
```

```{r}
outcomes <- c()
for (bmi in df_web$BMI) {
  if (bmi < 18.5) {
    outcomes <- c(outcomes, '0_undwerweight')
  } else if (bmi < 25) {
    outcomes <- c(outcomes, '1_normal')
  } else if (bmi < 30) {
    outcomes <- c(outcomes, '2_overweight')
  } else {
    outcomes <- c(outcomes, '3_obese')
  }
}


df_web$label <- outcomes
```

```{r}
####BMI labeling function####
create_bmi_label <- function(bmi){
  ifelse(bmi < 18.5, '0_underweight',
         ifelse(bmi < 25, '1_normal',
                ifelse(bmi < 30, '2_overweight', '3_obese')))
}
#The previous code for creating the label is a bit long and not very efficient, so I think it is better to use a function to create the label based on the BMI value. This will make the code more modular and easier to read.

df_web$label <- create_bmi_label(df_web$BMI)
```

```{r}
df_web[10:20, ]
```

```{r}
print(colnames(df))

df_web$MTRANS[df_web$MTRANS == 'Walking'] <- 1
df_web$MTRANS[df_web$MTRANS == 'Bike'] <- 1
df_web$MTRANS[df_web$MTRANS == 'Public_Transportation'] <- 0
df_web$MTRANS[df_web$MTRANS == 'Automobile'] <- 0
df_web$MTRANS[df_web$MTRANS == 'Motorbike'] <- 0
df_web$MTRANS <- as.numeric(df_web$MTRANS)

df_web$Gender[df_web$Gender == 'Male'] <- 1
df_web$Gender[df_web$Gender == 'Female'] <- 0
df_web$Gender <- as.numeric(df_web$Gender)

Xcol_binary <- c(
  'family_history_with_overweight',
  'FAVC',
  'SMOKE',
  'SCC'
)

for (col in Xcol_binary) {
  df_web[df_web[[col]] == 'yes', col] <- '1'
  df_web[df_web[[col]] == 'no', col] <- '0'
  df_web[[col]] <- as.numeric(df_web[[col]])
}

Xcol_ordinal <- c('CAEC', 'CALC')

df_web$CAEC <- as.numeric(factor(df_web$CAEC, levels = c('no', 'Sometimes', 'Frequently', 'Always'))) - 1
df_web$CALC <- as.numeric(factor(df_web$CALC, levels = c('no', 'Sometimes', 'Frequently', 'Always'))) - 1
```

```{r}
X_columns <- c('Age', 'Gender', 'family_history_with_overweight',
               'FAVC', 'FCVC', 'NCP', 'CAEC', 'SMOKE', 'CH2O', 'SCC', 'FAF', 'TUE',
               'CALC', 'MTRANS')
```

## Regression problem

```{r}
classes <- sort(unique(df_web$label))
freq <- table(df_web$label)
class_weight <- length(df_web$label) / (length(classes) * as.numeric(freq))
names(class_weight) <- names(freq)

print(classes)
print(class_weight)
```

```{r}
# Note: result can deviate from the python code due to random seed!
set.seed(42)
idx <- seq_len(nrow(df_web))

idx_train <- sample(idx, size = floor(0.7 * length(idx)))
idx_test <- setdiff(idx, idx_train)

X_train <- df_web[idx_train, X_columns]
X_test <- df_web[idx_test, X_columns]
y_train <- df_web$BMI[idx_train]
y_test <- df_web$BMI[idx_test]

sample_weights <- class_weight[df_web$label[idx_train]]
```

```{r}
####train test function####
train_test <- function(df, train_prop = 0.7, seed = 42) {
  set.seed(seed)
  idx <- seq_len(nrow(df))
  
  idx_train <- sample(idx, size = floor(train_prop * length(idx)))
  idx_test <- setdiff(idx, idx_train)
  
  list(train = idx_train, test = idx_test)
}

#The code of set.seed in the previouse code is a bit long, so I change it use the function to make it simple and easy to reuse
```


```{r}
reg <- lm(BMI ~ ., data = cbind(X_train, BMI = y_train))
reg <- randomForest(x = X_train, y = y_train, ntree = 100, nodesize = 5, maxnodes = 20)
```

```{r}
####train model function####
train_model <- function(X_train, y_train) {
  randomForest(x = X_train, y = y_train, ntree = 100, nodesize = 5, maxnodes = 20)
}

reg <- train_model(X_train, y_train)
```

```{r}
preds <- predict(reg, X_test)

plot(preds, y_test, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```

```{r}
plot(predict(reg, X_train), y_train, pch = 16, col = rgb(0, 0, 0, 0.5))
abline(a = 0, b = 1, col = 'red', lty = 3)
```


```{r}
sample_weights_test <- class_weight[df_web$label[idx_test]]

mae_baseline <- sum(abs(y_test - rep(mean(y_train), length(y_test))) * sample_weights_test) / sum(sample_weights_test)
mae_model <- mean(abs(y_test - preds) * sample_weights_test)

cat(sprintf('Mean absolute error (baseline): %.3f\n', mae_baseline))
cat(sprintf('Mean absolute error: %.3f\n', mae_model))
```

```{r}
renv::restore()
```

I apply function to avoiding code repetition (DRY) and make the code more modular and easier to understand and maintain.
I also need package 'renv' to manage the dependencies and make sure that the code can be reproduced in the future without worrying about package versions.

```{r}
source("src//download_data.R")
source("src//preprocess_data.R")
source("src//bmi_functions.R")
source("src//train_test.R")
source("src//train_model.R")
source("src//evaluate_model.R")
```

```{r}
library(mypackage)
```